### 2019-02-26 sections/abstract.tex
Added loads of fixme's based on Rinus and Alejandro's comments

% Context
Software that models how people work is omnipresent in today's society.
Current languages and frameworks are limited in flexibility and lack a good level of abstraction.
Task-oriented programming (TOP) is a programming paradigm that aims to provide the desired level of abstraction while still being expressive enough to describe real world collaboration.
It prescribes a declarative programming style to specify multi-user workflows.
Workflows can be higher-order.
They communicate through typed values on a local and global level.
Such specifications can be turned into interactive applications for different platforms, supporting collaboration during execution.

% Inquiry
TOP has been around for more than a decade, in the form of iTasks, which is tailored for real-world usability.
So far, it has not been given a formalization in the style common in programming language research.

>>>>

% Context
Software that models how people work is omnipresent in today's society.
Current languages and frameworks are limited in flexibility and lack a good level of abstraction. \fixme{Rinus: het is vast waar, maar ik zou het iets zachter formuleren: focus op een concreet toepassingsgebied, maar minder op ...}
Task-oriented programming (TOP) is a programming paradigm that aims to provide the desired level of abstraction while still being expressive enough to describe real world collaboration.
It prescribes a declarative programming style to specify multi-user workflows.
Workflows can be higher-order.
They communicate through typed values on a local and global level.
Such specifications can be turned into interactive applications for different platforms, supporting collaboration during execution.

% Inquiry
TOP has been around for more than a decade, in the form of iTasks, which is tailored for real-world usability.
So far, it has not been given a formalization in the style common in programming language research. \fixme{Rinus: not been given a formalization which is very suitable for formal reasoning.}

### 2019-02-26 sections/example.tex
scale figures for new format
compact code examples
make figure bigger
move figure in example
Added loads of fixme's based on Rinus and Alejandro's comments

The requirements of the application are as follows.
\begin{enumerate*}
  \item A user has to enter a list of passengers for which to book tickets.
  \item At least one of these passengers has to be an adult.
  \item After a valid list of passengers has been entered, the user has to pick seats.
  \item Only free seats may be picked.
  \item Every passenger must have exactly one seat.
  \item Multiple users should be able to book tickets at the same time.
\end{enumerate*}
For this example we assume that the host language has lists and four functions on them: all, any, intersect, and difference.
%%NOTE: After Peter's comments symbols are not clear --TS
We also make use of string equality ($\equiv$), dereferencing ($!$), reference assignment ($:=$), and expression sequencing ($;$).
For brevity, we omit the type annotations of variable bindings.

@@

Now we develop our workflow in a top down way.
A flight booking starts with an interactive task \TS{enter (List Passenger)}
where users can enter a list of passengers.
A passenger is valid if the name is not empty and the age is at least 0.
A list of passengers is valid if each passenger is valid, and at least one of the passengers is an adult.
When the user has entered a valid list of passengers, the step after \TS{>>?} becomes enabled,
and the user can proceed to picking seats.
\begin{TASK}
  let valid = \p. not (fst p == "") /\ snd p >= 0 in
  let adult = \p. snd p >= 18 in
  let allValid = \ps. all valid ps /\ any adult ps in
  let bookFlight = enter (List Passenger) >>? \ps.
    if allValid ps then chooseSeats ps else fail
\end{TASK}
A selection of seats is correct if every entered seat is free.
\begin{TASK}
  let correct = \ss. intersect ss !freeSeats == ss in
  let chooseSeats = \ps. enter (List Seat) >>? \ss.
    if correct ss /\ length ps == length ss
      then confirmBooking ps ss else fail
\end{TASK}
The function \TS{confirmBooking} removes the picked seats from the shared list of free seats,
and displays the end result using the \TS{edit}-construct.
\begin{TASK}
  let confirmBooking = \ps. \ss.
    freeSeats := difference !freeSeats ss; edit <<ps, ss>>
\end{TASK}
% It uses a function \TS{difference}, which removes all elements from the second list from the first list.

@@

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{figures/flight-booking.png}
  \caption{
    Running web application of the flight booking example using a translation to \ITASKS.
    It shows three users booking a flight simultaneously.
    The first user entered name and age and continued picking seats.
    The second is entering details of two passengers.
    The ages are not filled in, therefore the \TS{Continue} button is disabled.
    The third user finished a booking.
    The first user can not pick seats \smallcaps{1b} and \smallcaps{1c} any more.
    The message bubble shows that the \TS{age} field only accepts integer values.
  }
  \label{fig:flight-booking}
\end{figure}

>>>>

The requirements of the application are as follows.
\begin{enumerate*}
  \item A user has to enter a list of passengers for which to book tickets.
  \item At least one of these passengers has to be an adult.
  \item After a valid list of passengers has been entered, the user has to pick seats.
  \item Only free seats may be picked.
  \item Every passenger must have exactly one seat.
  \item Multiple users should be able to book tickets at the same time.
\end{enumerate*}
For this example we assume that the host language has lists and four functions on them: all, any, intersect, and difference. \fixme{Alejandro: small description may be needed} %nico: of the functions
%%NOTE: After Peter's comments symbols are not clear --TS
We also make use of string equality ($\equiv$), dereferencing ($!$), reference assignment ($:=$), and expression sequencing ($;$).
For brevity, we omit the type annotations of variable bindings.

\begin{figure}[b]
  \includegraphics[width=0.8\columnwidth]{figures/flight-booking.png}
  \caption{
    Running web application of the flight booking example using a translation to \ITASKS.
    It shows three users booking a flight simultaneously.
    The first user entered name and age and continued picking seats.
    The second is entering details of two passengers.
    The ages are not filled in, therefore the \TS{Continue} button is disabled.
    The third user finished a booking.
    The first user can not pick seats \smallcaps{1b} and \smallcaps{1c} any more.
    The message bubble shows that the \TS{age} field only accepts integer values.
  }
  \label{fig:flight-booking}
\end{figure}

@@

Now we develop our workflow in a top down way.
A flight booking starts with an interactive task \TS{enter (List Passenger)}
where users can enter a list of passengers.
A passenger is valid if the name is not empty and the age is at least 0.
A list of passengers is valid if each passenger is valid, and at least one of the passengers is an adult.
When the user has entered a valid list of passengers, the step after \TS{>>?} becomes enabled,
and the user can proceed to picking seats.
\begin{TASK}
  let valid = \p. not (fst p == "") /\ snd p >= 0 in
  let adult = \p. snd p >= 18 in
  let allValid = \ps. all valid ps /\ any adult ps in
  let bookFlight = enter (List Passenger) >>? \ps.
    if allValid ps then chooseSeats ps else fail
\end{TASK}
\fixme{Alejandro: At this point I wonder, what is $\Enter$}
A selection of seats is correct if every entered seat is free.
\begin{TASK}
  let correct = \ss. intersect ss !freeSeats == ss in
  let chooseSeats = \ps. enter (List Seat) >>? \ss.
    if correct ss /\ length ps == length ss then confirmBooking ps ss else fail
\end{TASK}

The function \TS{confirmBooking} removes the picked seats from the shared list of free seats,
and displays the end result using the \TS{edit}-construct.\fixme{Alejandro: Give \TS{edit} a proper name?}
\begin{TASK}
  let confirmBooking = \ps. \ss.
    freeSeats := difference !freeSeats ss; edit <<ps, ss>>
\end{TASK}
% It uses a function \TS{difference}, which removes all elements from the second list from the first list.

@@

### 2019-02-26 sections/introduction.tex
Added loads of fixme's based on Rinus and Alejandro's comments

@@

\subsection{Task-oriented programming}

% Task-Oriented Programming (\TOP) is a programming paradigm aimed at writing interactive multi-user applications in a declarative way \cite{conf/ppdp/PlasmeijerLMAK12}.
Task-oriented programming (\TOP) is a programming paradigm which aims to be the sweet spot between faithful modelling workflows
and rapid prototyping of multi-user web applications supporting these workflows.
% Task-oriented programming (\TOP) is a programming paradigm to support these ways of working.
\TOP focusses on modelling collaboration patterns.
This gives rise to user's need to interact and share information.
Next to that, \TOP automatically provides solutions to common development jobs like designing \GUI\ s, connecting to databases, and communicating between servers and clients.

Therefore,
a language that supports \TOP should choose the right level of abstraction to support two things.
% \TOP has two aspects.
% First, it should allow to specify tasks from real world scenarios.
Firstly, it should provide primitive building blocks that are useful for high-level descriptions of how users collaborate with each other and with machines.
These building blocks are \emph{editors}, \emph{composition}, and \emph{shared data}.
% Second, it should be able to generate multi-user web applications to support these scenarios.
Secondly, it should be able to generate applications, including graphical user interfaces, from workflows modelled with said building blocks.

@@

Currently, we know of two frameworks using \TOP: \ITASKS and \MTASKS.

@@

Both implementations currently lack formal semantics which can be used to prove properties about tasks.

@@

One of the defining aspects of \ITASKS is its embedding in a functional programming language.
This creates a synergy where \ITASKS profits from the expressivity of Clean, while enhancing the language with functionality for user interaction.
The tight integration of task DSL and host language makes it sometimes difficult to see where the boundaries are.

>>>>

\fixme{Alejandro: The introduction seems to be more about why Task-Oriented Programming
is nice and useful. Personally, I would like it to tell me another
story: given that some people find TOP useful (as shown by the many
examples), why should be care about formalizing it and which challenges
does it pose? I think this is what people from ICFP would care about more.}

@@

\subsection{Task-oriented programming}
\fixme{Alejandro: Not entirely convincing}%Nico: hij heeft het hier volgens mij over de titel van de sectie en het woord aims in de eerste zin. Misschien moeten we een hardere claim doen hier?
% Task-Oriented Programming (\TOP) is a programming paradigm aimed at writing interactive multi-user applications in a declarative way \cite{conf/ppdp/PlasmeijerLMAK12}.
Task-oriented programming (\TOP) is a programming paradigm which aims to be the sweet spot between faithful modelling workflows
and rapid prototyping of multi-user web applications supporting these workflows.\fixme{Rinus: Hier zou ik een referentie naar ons ppdp2012 paper verwachten.}
% Task-oriented programming (\TOP) is a programming paradigm to support these ways of working.
\TOP focusses on modelling collaboration patterns.
This gives rise to user's need to interact and share information.
Next to that, \TOP automatically provides solutions to common development jobs like designing \GUI\ s, connecting to databases, and communicating between servers and clients.

Therefore,
a language that supports \TOP should choose the right level of abstraction to support two things.
% \TOP has two aspects.
% First, it should allow to specify tasks from real world scenarios.
Firstly, it should provide primitive building blocks that are useful for high-level descriptions of how users collaborate with each other and with machines.
These building blocks are: \emph{editors}, \emph{composition}, and \emph{shared data}.
% Second, it should be able to generate multi-user web applications to support these scenarios.
Secondly, it should be able to generate applications, including graphical user interfaces, from workflows modelled with said building blocks.

@@

Currently, we know of two frameworks implementing \TOP: \ITASKS and \MTASKS.

@@

Both implementations currently lack formal semantics which can be used to prove properties about tasks.\fixme{Rinus: can be used.. -> is suited}

@@

One of the defining aspects of \ITASKS is its embedding in a functional programming language.
This creates a synergy where \ITASKS profits from the expressivity of Clean, while enhancing the language with functionality for user interaction.
The tight integration of task DSL and a host language makes it sometimes difficult to see where the boundaries are.

### 2019-02-26 sections/intuition.tex
Added loads of fixme's based on Rinus and Alejandro's comments

The way the combinators are defined matches real life closely.
When we want to have breakfast, we have to complete several other tasks first before we can do so.
We decide what we want to have and then prepare it.
We can prepare the different items we have for breakfast in parallel, but not at the same time.
For example, it is impossible to scramble eggs, and put on the kettle for tea simultaneously.
Instead, what is meant by parallel is that the order in which we do tasks and the smaller tasks that they are composed of, does not matter.
Then finally, only when every item we want to have for breakfast is ready, can we sit down and enjoy it.
\end{example}

@@

In \TOPHAT, editors are the basic method of communication with the environment.
There are different editors, denoted by different box symbols.
An editor of the form $\Edit v$, where $v$ is of type $\tau$, reacts to two kinds of events:
Change events $v$ with a value of type $\tau$ and clear events $\Empty$ which empty the editor.
This results in a typed empty editor $\Enter \tau$, which is annotated with the same type.

@@

@@

Several observations can be made about tasks.
One of those is determining the value of a task $\Value(t)$.
Not all tasks have a value, for example the empty editor $\Enter \tau$, which makes $\Value$ a partial function.

@@

The final observation is to determine whether a task results in $\Fail$ (failure).
The step combinator and the choice combinator use this to prevent users from picking a failing task.

@@

Let $t_1$ and $t_2$ be tasks.
The parallel combination $t_1 \And t_2$ stands for two independent tasks carried out at the same time.
Therefore $\And$ introduces concurrency.
For the system it does not matter if the tasks are executed by two people actually in parallel, or by one person who switches between the tasks.
The inputs sent to the component tasks are interleaved into a serial stream, which is sent to the parallel combinator.
We assume that such a serialization is always possible.
The tasks are truly independent of each other, all interleavings are permitted.
The environment must prefix events to $t_1$ and $t_2$ by $\First$ (first) and $\Second$ (second) respectively.
This unambiguously renames the inputs, removing any source of nondeterminism.

@@

To model collaboration, one needs communication, concurrency, and synchronisation.
In \TOP, collaboration is specified on a high level of abstraction, hiding the details of communication.
For example, the cigarette smokers communicate with each other, but the programs do not explicitly mention sending or receiving events.

>>>>

The way the combinators are defined matches real life closely.
When we want to have breakfast, we have to complete several other tasks first before we can do so.
We decide what we want to have and then prepare it.
We can prepare the different items we have for breakfast in parallel, but not at the same time.
For example, it is impossible to scramble eggs, and put on the kettle for tea simultaneously.
Instead, what is meant by parallel is that the order in which we do tasks and the smaller tasks that they are composed of, does not matter.
\fixme{Alejandro has placed a line and exclamation under this sentence. I'm not sure what to do here.}
Then finally, only when every item we want to have for breakfast is ready, can we sit down and enjoy it.
\end{example}

@@

In \TOPHAT, editors are the basic method of communication with the environment.
There are different editors, denoted by different box symbols.
An editor of the form $\Edit v$, where $v$ is of type $\tau$, reacts to two kinds of events:
Change events $v$ with a value of type $\tau$ and clear events $\Empty$ which empty the editor.
This results in a typed empty editor $\Enter \tau$, which is annotated with the same type.
\fixme{Alejandro: unclear} %Nico: He has marked the above paragraph as being unclear.

@@

\fixme{Alejandro: What is a value? What is a type?}

@@

Several observations can be made about tasks.
One of those is determining the value of a task $\Value(t)$. \fixme{Should we change the type of Value to Maybe v? Both johan and alejandro mention that it is not completely clear to them.}
Not all tasks have a value, for example the empty editor $\Enter \tau$, which makes $\Value$ a partial function.

@@

The final observation is to determine whether a task results in $\Fail$ (failure).
The step combinator $\Next$ and the choice combinator $\Xor$ use this to prevent users from picking a failing task.

@@

Let $t_1$ and $t_2$ be tasks.
The parallel combination $t_1 \And t_2$ stands for two independent tasks carried out at the same time.
Therefore $\And$ introduces concurrency. \fixme{Alejandro: But before you said it was interleaving!}%nico: en later doen we dit ook.
For the system it does not matter if the tasks are executed by two people actually in parallel, or by one person who switches between the tasks.
The inputs sent to the component tasks are interleaved into a serial stream, which is sent to the parallel combinator.
We assume that such a serialization is always possible.
The tasks are truly independent of each other, all interleavings are permitted.
The environment must prefix events to $t_1$ and $t_2$ by $\First$ (first) and $\Second$ (second) respectively.
This unambiguously renames the inputs, removing any source of nondeterminism.

@@

To model collaboration, one needs communication, concurrency, and synchronisation. \fixme{Alejandro: Is this a basic calculus to why iTasks is elaborated?} %nico: not sure what he means here...
In \TOP, collaboration is specified on a high level of abstraction, hiding the details of communication.
For example, the cigarette smokers communicate with each other, but the programs do not explicitly mention sending or receiving events.

### 2019-02-26 sections/language.tex
added fixme about beta
Added loads of fixme's based on Rinus and Alejandro's comments

\label{sub:expressions}
The host language is a simply typed lambda calculus, extended with some basic types and \ML-style references.
The grammar in \cref{fig:language-grammar} defines the syntax of the host language.
It has abstractions, applications, variables, and constants for booleans, integers and strings.
The symbol $\star$ stands for binary operators.
For the result of parallel tasks we need pairs.
Conditionals come in handy for defining guards.
%
References will be used to implement shared editors.
Our treatment of references closely follows the one by \citet{books/Pierce02TAPL}.
Creating a reference using the keyword $\Ref$ yields a location $l$.
Locations are not intended to be directly manipulated by the programmer.
The symbols ! and $:=$ stand for dereferencing and assignment.
The unit value will be used as the result of assignments.

\begin{figure}[h]
  \small
  \usemacro{G-Language-Compact}
  \caption{Language grammar} \label{fig:language-grammar}
\end{figure}

\label{sub:notation}
We use double quotation marks to denote strings.
Integers are denoted by their decimal representation, and booleans are written $\True$ and $\False$.
We freely make use of the logic operators $\Not$, $\land$, and $\lor$, arithmetic operators $+$, $-$, $\times$, and the string append operator $\pp$.
Furthermore, we use standard comparison operations $<$, $\le$, $\equiv$, $\not\equiv$, $\ge$, and $>$.
The symbol $\star$ stands for any of those.
%
\label{sub:abbreviations}
The notation $e_1; e_2$ is an abbreviation for $(\lambda x:\Unit.\ e_2)\ e_1$, where $x$ is a fresh variable.
The notation $\Let x:\tau = e_1 \In e_2$ is an abbreviation for $(\lambda x:\tau.\ e_2)\ e_1$.

\label{sub:pretasks}
The grammar in \cref{fig:task-grammar} specifies the syntactic category of \emph{pretasks}.
Pretasks are tasks that have unevaluated subexpressions.
Each pretask will be discussed in more detail in the following subsections.
We use open symbols ($\Edit, \Enter, \Next, \Xor$) for tasks that require user input, and closed symbols ($\Update, \Then, \Or$) for tasks that can be evaluated without user input.

@@

Typing rules are of the form $\RelationT$, which should be read as \enquote{in environment $\Gamma$ and store typing $\Sigma$, expression $e$ has type $\tau$}.
Typing rules for expressions in the host language are presented in the appendix.
The typing rules for pretasks are given in \cref{fig:typing-rules}.

\begin{figure}[h]
  \small
  \begin{gather*}
    \boxed{\RelationT} \Break
    \userule{T-Edit} \Quad
    \userule{T-Enter} \Quad
    \userule{T-Update} \Quad
    \userule{T-Fail} \Break
    \userule{T-Then} \Quad
    \userule{T-Next} \Break
    \userule{T-And} \Quad
    \userule{T-Or} \Quad
    \userule{T-Xor}
  \end{gather*}
  \caption{Typing rules} \label{fig:typing-rules}
\end{figure}

@@

\paragraph{Internal and external step $(t \Then e, t \Next e)$}
\label{sub:steps}

@@

External choice ($\Xor$) is different in this regard.
An external choice requires users to pick a branch before continuing with it.
This means users cannot work on the branches of an external choice before picking one.

>>>>

\label{sub:expressions}
The host language is a simply typed lambda calculus, extended with some basic types and \ML-style references.\fixme{Alejandro: why references?}
The grammar in \cref{fig:language-grammar} defines the syntax of the host language.
It has abstractions, applications, variables, and constants for booleans, integers and strings.
The symbol $\star$ stands for binary operators.
For the result of parallel tasks we need pairs.
Conditionals come in handy for defining guards.
%
References will be used to implement shared editors.
Our treatment of references closely follows the one by \citet{books/Pierce02TAPL}.
Creating a reference using the keyword $\Ref$ yields a location $l$.
Locations are not intended to be directly manipulated by the programmer.
The symbols ! and $:=$ stand for dereferencing and assignment.
The unit value will be used as the result of assignments.

\begin{figure}[h]
  \small
  \usemacro{G-Language-Compact}
  \caption{Language grammar} \label{fig:language-grammar}
\end{figure}
\fixme{Alejandro about l: different syntax from vars?}
\label{sub:notation}
We use double quotation marks to denote strings.
Integers are denoted by their decimal representation, and booleans are written $\True$ and $\False$.
We freely make use of the logic operators $\Not$, $\land$, and $\lor$, arithmetic operators $+$, $-$, $\times$, and the string append operator $\pp$.
Furthermore, we use standard comparison operations $<$, $\le$, $\equiv$, $\not\equiv$, $\ge$, and $>$.
The symbol $\star$ stands for any of those.
%
\label{sub:abbreviations}
The notation $e_1; e_2$ is an abbreviation for $(\lambda x:\Unit.\ e_2)\ e_1$, where $x$ is a fresh variable.
The notation $\Let x:\tau = e_1 \In e_2$ is an abbreviation for $(\lambda x:\tau.\ e_2)\ e_1$.

\label{sub:pretasks}
The grammar in \cref{fig:task-grammar} specifies the syntactic category of \emph{pretasks}.
Pretasks are tasks that have unevaluated subexpressions. \fixme{Alejandro: How do you define unevaluated?}
Each pretask will be discussed in more detail in the following subsections.
We use open symbols ($\Edit, \Enter, \Next, \Xor$) for tasks that require user input, and closed symbols ($\Update, \Then, \Or$) for tasks that can be evaluated without user input.

@@

Typing rules are of the form $\RelationT$, which should be read as \enquote{in environment $\Gamma$ and store typing $\Sigma$, expression $e$ has type $\tau$}.
Typing rules for expressions in the host language are presented in the appendix.
The typing rules for pretasks are given in \cref{fig:typing-rules}.
\fixme{Say someting about why we need beta, and use beta in the acutal typing rule}
\fixme{Alejandro: Is there anything surprising? Say what}
\begin{figure}[h]
  \small
  \begin{gather*}
    \boxed{\RelationT} \Break
    \userule{T-Edit} \Quad
    \userule{T-Enter} \Quad
    \userule{T-Update} \Quad
    \userule{T-Fail} \Break
    \userule{T-Then} \Quad
    \userule{T-Next} \Break
    \userule{T-And} \Quad
    \userule{T-Or} \Quad
    \userule{T-Xor}
  \end{gather*}
  \caption{Typing rules} \label{fig:typing-rules}
\end{figure}

@@

\paragraph{Internal and external step $(t \Then e, t \Next e)$}\fixme{Alejandro: The naming is confusing to me}
\label{sub:steps}

@@

External choice ($\Xor$) is different in this regard.
An external choice requires users to pick a branch before continuing with it.
This means users cannot work on the branches of an external choice before picking one. \fixme{Alejandro: Is this actually implied?}

### 2019-02-26 sections/properties.tex
Added loads of fixme's based on Rinus and Alejandro's comments

Additionally, we show that the normalisation semantics is a big-step semantics.
While at first sight, this might seem obvious, the fact that we are dealing with
state complicates matters.

@@

\subsection{Preservation}
\label{sub:preservation}

@@

\begin{theorem}[Preservation under evaluation]
  % For all well typed expressions $e$ and states $s$,
  For all expressions $e$ and states $s$
  such that $\Gamma,\Sigma \infers e:\tau$ and $\Gamma,\Sigma \infers s$,
  if $e,s \evaluate e',s'$,
  then $\Gamma,\Sigma \infers e':\tau$ and $\Sigma \infers s'$.
  \label{thm:pres-eval}
\end{theorem}

@@

\begin{theorem}[Preservation under normalisation]
  % For all well typed expressions $e$ and states $s$,
  For all expressions $e$ and states $s$
  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
  if $e,s \normalise e',s'$,
  then $\Gamma,\Sigma \infers e':\Task\tau$ and $\Sigma \infers s'$.
  \label{thm:pres-norm}
\end{theorem}

@@

\begin{lemma}[Task value preserves]
  % For all well typed expressions $e$ and states $s$,
  For all expressions $e$ and states $s$
  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
  if $\Value{(e,s)}=v$,
  then $v:\tau$.
  \label{lem:presvalue}
\end{lemma}

\begin{lemma}[Striding preserves]
  For all expressions $e$ and states $s$
  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
  if $e,s \stride e',s'$,
  then $\Gamma,\Sigma \infers e':\Task\tau$ and $\Sigma \infers s'$.
  \label{thm:pres-stride}
\end{lemma}

@@

\begin{theorem}[Preservation under handling]
  % For all well typed expressions $e$, states $s$, and inputs $i$,
  For all expressions $e$, states $s$ and inputs $i$
  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
  if $ e,s \handle{i} e',s'$,
  then $\Gamma,\Sigma\infers e':\Task\tau$ and $\Sigma\infers s'$.
  \label{thm:pres-handle}
\end{theorem}

>>>>

Additionally, we show that the normalisation semantics is a big-step semantics. \fixme{Alejandro: ?}
While at first sight, this might seem obvious, the fact that we are dealing with
state complicates matters.

@@

\subsection{Type preservation}
\label{sub:preservation}

@@

\begin{theorem}[Type preservation under evaluation]
  % For all well typed expressions $e$ and states $s$,
  For all expressions $e$ and states $s$
  such that $\Gamma,\Sigma \infers e:\tau$ and $\Gamma,\Sigma \infers s$,
  if $e,s \evaluate e',s'$,
  then $\Gamma,\Sigma \infers e':\tau$ and $\Sigma \infers s'$.
  \label{thm:pres-eval}
\end{theorem}

@@

\begin{theorem}[Type preservation under normalisation]
  % For all well typed expressions $e$ and states $s$,
  For all expressions $e$ and states $s$
  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
  if $e,s \normalise e',s'$,
  then $\Gamma,\Sigma \infers e':\Task\tau$ and $\Sigma \infers s'$.
  \label{thm:pres-norm}
\end{theorem}

@@

\begin{lemma}[Task value preserves types]
  % For all well typed expressions $e$ and states $s$,
  For all expressions $e$ and states $s$
  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
  if $\Value{(e,s)}=v$,
  then $v:\tau$.
  \label{lem:presvalue}
\end{lemma}

\begin{lemma}[Striding preserves types]
  For all expressions $e$ and states $s$
  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
  if $e,s \stride e',s'$,
  then $\Gamma,\Sigma \infers e':\Task\tau$ and $\Sigma \infers s'$.
  \label{thm:pres-stride}
\end{lemma}

@@

\begin{theorem}[Type preservation under handling]
  % For all well typed expressions $e$, states $s$, and inputs $i$,
  For all expressions $e$, states $s$ and inputs $i$
  such that $\Gamma,\Sigma \infers e:\Task\tau$ and $\Sigma \infers s$,
  if $ e,s \handle{i} e',s'$,
  then $\Gamma,\Sigma\infers e':\Task\tau$ and $\Sigma\infers s'$.
  \label{thm:pres-handle}
\end{theorem}

### 2019-02-26 sections/relatedwork.tex
Added loads of fixme's based on Rinus and Alejandro's comments

\paragraph{iTasks}

As mentioned earlier, iTasks is an implementation of \TOP.
iTasks has many features, and its basic combinators are versatile and powerful.
Simpler combinators are implemented by restricting the powerful ones.
This is useful for everyday programming, where having lots of functionality at one's fingertips is convenient and efficient.

@@

>>>>

\paragraph{iTasks}
\fixme{Rinus: Waarom geen korte opsomming van iTask features die niet zijn meegenomen in TopHat?}
As mentioned earlier, iTasks is an implementation of \TOP.
iTasks has many features, and its basic combinators are versatile and powerful.
Simpler combinators are implemented by restricting the powerful ones.
This is useful for everyday programming, where having lots of functionality at one's fingertips is convenient and efficient.

@@

\fixme{Rinus: Volgens mij is er wel degelijk een formele semantiek, omdat heel iTask wordt uitgelegd in termen van standaard Clean functies,
en Clean heeft een semantiek (graaf herschrijf systeem).
Dat de beschrijving onhandig (wellicht te low level is) is voor het formeel redeneren over iTask programma's is een beter argument.}
\fixme{Alejandro: What is missing from \TOPHAT to describe i or m Tasks}

### 2019-02-26 sections/semantics.tex
changed UI fixme to comment
All changes based on johan's comments
Added loads of fixme's based on Rinus and Alejandro's comments

@@

\begin{figure}[h]
  \small
  \usemacro{G-Values-Compact}
  \caption{Value grammar} \label{fig:value-grammar}
\end{figure}

The rules to evaluate expressions $e$ do not differ from standard work, except for the task constructs.
The evaluation rules for tasks can be deduced from the value grammar.
They are given in the appendix.

@@

Steps whose left hand sides are failing can never proceed because of the lack of an observable value.
Therefore they are itself failing.
The internal choice of two failing tasks is failing.
External choices let the user pick a side and only then evaluate the corresponding subexpression.
To determine if an external choice is failing, it needs to peek into the future to check if both subexpressions are failing.
Annotations are failing only if the inner task is failing.

@@

\fixme{Add $\UserInterface$ to appendix and refer.}
\TOPHAT is designed such that a user interface can be generated from a task's syntax tree.
A possible graphical user interface is shown in \cref{fig:flight-booking}, where tasks are rendered as \HTML pages.
Editors are rendered as input fields,
external choices are represented by two buttons,
and parallel tasks are rendered side by side.
Steps only show the interface of their left hand side.
In case of an external step they are accompanied by a button.
When the guard condition of a step is not fulfilled, the button is disabled.

@@

The split between striding and normalisation is due to mutable references.
Consider the following example, where $s=\set{l\mapsto\False}$.
\begin{TASK}
  (update l >>= \x:Bool. if x then e else fail) <&> (l := True; edit <<>>)
\end{TASK}
\refrule{S-And} reduces this expression in one step to
\begin{TASK}
  (update l >>= \x:Bool. if x then e else fail) <&> (edit <<>>)
\end{TASK}
with $s'=\set{l\mapsto\True}$.
This expression is not normalised, because the left task can take a step.
The issue here lies in the fact that the right task updates $l$.
To overcome this problem, the \refrule{N-Done} and \refrule{N-Repeat} rules ensure that striding is applied until the state $s$ becomes stable and no further normalisation can take place.

@@

\begin{figure}[h]
  \small
  \usemacro{G-Inputs-Compact}
  \caption{Input grammar} \label{fig:input-grammar}
\end{figure}

\begin{figure}[h]
  \small
  \usemacro{O-Inputs}
  \caption{Inputs} \label{fig:observation-input}
\end{figure}

>>>>

\fisme{Alejandro: In the formalization you show quite some figures. I think that for
each figure you need to be more explicit in the text about what is
unsurprising and what is new or should get the attention of the reader.}

@@

\begin{figure}[h]
  \small
  \usemacro{G-Values-Compact}
  \caption{Value grammar} \label{fig:value-grammar}
\end{figure}
\fixme{Alejandro over tasks: I was not expecting these to be values...}
The rules to evaluate expressions $e$ do not differ from standard work, except for the task constructs.
The evaluation rules for tasks can be deduced from the value grammar.
They are given in the appendix. \fixme{Alejandro: What is interesting about them?}

@@

Steps whose left hand sides are failing can never proceed because of the lack of an observable value.
Therefore they are itself failing.
The internal choice of two failing tasks is failing.
External choices let the user pick a side and only then evaluate the corresponding subexpression.
To determine if an external choice is failing, it needs to peek into the future to check if both subexpressions are failing.

@@

%\fixme{Add $\UserInterface$ to appendix and refer.}
\TOPHAT is designed such that a user interface can be generated from a task's syntax tree.
A possible graphical user interface is shown in \cref{fig:flight-booking}, where tasks are rendered as \HTML pages.
Editors are rendered as input fields,
external choices are represented by two buttons,
and parallel tasks are rendered side by side.
Steps only show the interface of their left hand side.
In case of an external step they are accompanied by a button.
When the guard condition of a step is not fulfilled, the button is disabled.

@@

The split between striding and normalisation is due to mutable references.
Consider the following example, where $s=\set{l\mapsto\False}$.
\begin{TASK}
  (update l >>= \x:Bool. if x then e else fail) <&> (l := True; edit <<>>)
\end{TASK}
\refrule{S-And} reduces this expression in one step to
\begin{TASK}
  (update l >>= \x:Bool. if x then e else fail) <&> (edit <<>>)
\end{TASK}
with $s'=\set{l\mapsto\True}$.
This expression is not normalised, because the left task can take a step.
The issue here lies in the fact that the right task updates $l$.
To overcome this problem, the \refrule{N-Done} and \refrule{N-Repeat} rules ensure that striding is applied until the state $s$ becomes stable and no further normalisation can take place.
\fixme{The example is not 100\% clear to me}

@@

\begin{figure}[h]
  \small
  \usemacro{G-Inputs-Compact}
  \caption{Input grammar} \label{fig:input-grammar}
\end{figure}
\fixme{Fi, Si are sort of a path to the task}
\begin{figure}[h]
  \small
  \usemacro{O-Inputs}
  \caption{Inputs} \label{fig:observation-input}
\end{figure}
